Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-06-15 manuscript/manuscript.Rmd
adds QBM loyo simulation results and updates ms Table 1 with those results

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
   - \setlength\parindent{24pt}
bibliography: ~/Dropbox/Bibliography/MicroMeso.bib
csl: components/ecology.csl

@@

@@

##  QBM results
# Read in data
qbm_onestep <- readRDS("../analysis/quadBM/simulations/validation_results/qbm_one-step_forecasts_combined.rds")
# Average predictions over quad-year reps
avg_qbm <- ddply(qbm_onestep, .(species, quad, year), summarise,
                 avg_prediction = median(predcov),
                 observation = mean(obscov),
                 quant_dist = quantile(predcov*100, probs = 0.95) - quantile(predcov, probs = 0.05))
# Calculate error
avg_qbm$error <- avg_qbm$observation*100 - avg_qbm$avg_prediction*100
# Aggregate over species
qbmstats <- ddply(avg_qbm, .(species), summarise,
                  mae = mean(abs(error)),
                  mean_cover = mean(observation*100),
                  quant_dist = mean(quant_dist))
qbmstats$model <- "QBM"

mae_list <- list()
mae_pvals <- list()
species <- unique(qbmstats$species)
for(spp in species){
  tmp1 <- subset(tmpipm, species==spp)
  tmp2 <- subset(tmpqbm, species==spp)
  err1 <- as.data.frame(abs(tmp1$mean_obs - tmp1$mean_pred))
  err2 <- as.data.frame(abs(tmp2$mean_obs - tmp2$mean_pred))
  colnames(err1) <- "resid"
  err1$model <- "ipm"
  colnames(err2) <- "resid"
  err2$model <- "qbm"
  errd <-  rbind(err1, err2)
  mae_ttest <- with(errd, t.test(resid~model, alternative = "less"))
  mae_list[[spp]] <- mae_ttest
  mae_pvals[[spp]] <- mae_ttest[["p.value"]]
}

mae_ps <- melt(mae_pvals)

@@

\pagebreak{}
```{r table_2,results='asis',message=FALSE, comment=NA, echo=FALSE}
qbm_diffs <- readRDS("../montana/quadBM/simulations/results/qbm_climatesims_logdiffs.RDS")
qbm_diffs$type <- rep("QBM", nrow(qbm_diffs))
ipm_diffs <- readRDS("../montana/ipm/simulations/results/ipm_climatesims_logdiffs.RDS")
ipm_diffs$type <- rep("IPM", nrow(ipm_diffs))

diff_df <- rbind(qbm_diffs, ipm_diffs)
med_cover_perc <- as.data.frame(diff_df$med_cover*100)
med_cover_perc$model <- diff_df$type 
med_cover_perc$sim <- diff_df$variable
med_cover_perc$species <- diff_df$species
colnames(med_cover_perc)[1] <- "cover"
medcast <- dcast(med_cover_perc, species+sim~model, value.var = "cover")
colnames(medcast) <- c("Species", "Simulation", "IPM Change (%)", "QBM Change (%)")
simnames <- c("ppt", "temp", "ppt+temp")
medcast$Simulation <- rep(simnames, 4)
print.xtable(xtable(medcast, caption = "Percent change from equilibrium (temporal median of simulated plant cover) at observed climate relative to a 1\\% climate change as forecast by each model type."),type="latex", comment=FALSE,
             include.rownames=FALSE, caption.placement="top")
```

@@

Figures
-------
```{r fig.cap="Work flow of the data aggregation, model fitting, and population simulating.", fig.height=6}
library(png)
library(grid)
img <- readPNG("components/figure/micromeso_flow.png")
grid.raster(img)
```

@@

```{r figure_1, dependson="plot-options", fig.cap="Sensitivity of equilibrium cover to a 1% increase in precipitation (+ppt), temperature (+temp), or both (+ppt&temp) applied to individual and combined vital rates. For example, the points associated with "G" show the median cover from IPM simulations where a climate perturbation is applied only to the growth regression climate covariates.", fig.width=4.5, fig.height=6, cache=FALSE, results='hide'}
setwd("../montana/ipm/simulations/sensitivity_results/")
files <- list.files()
cover_files <- files[grep("cover", files)]
spp_id <- substr(cover_files, 1, 4)
num_files <- length(cover_files)
all_sims <- data.frame(time=NA, cover=NA, species=NA, climsim=NA, vital=NA)
for(i in 1:num_files){
  tmp <- read.csv(cover_files[i])
  tmp$species <- rep(spp_id[i], nrow(tmp))
  all_sims <- rbind(all_sims, tmp)
}
all_sims <- all_sims[2:nrow(all_sims),]

@@

equilibrium_cover <- subset(equilibrium_cover, vital!="A")
bothid <- which(equilibrium_cover[,"climsim"]=="ppttemp")
equilibrium_cover[bothid,"climsim"] <- "zppttemp"
# myCols2 <- c("#277BA8", "#7ABBBD", "#AED77A")
ggplot(equilibrium_cover, aes(x=climsim, y=eq_cover*100, 
                              shape=vital, group=vital, linetype=vital))+
  geom_hline(data=mean_cover, aes(yintercept=value*100), linetype=2, color="grey45")+
  geom_line()+
  geom_point(size=4)+
  facet_grid(species~., scales = "free")+
  scale_linetype_discrete(name="Vital Rate")+
  scale_shape_discrete(name="Vital Rate")+
  scale_x_discrete(labels=c("baseline", "+ppt", "+temp", "+ppt&temp"))+
  ylab("Equilibrium Cover (%)")+
  xlab("Simulation")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
setwd("~/Repos/MicroMesoForecast/manuscript")
```

```{r figure_2, dependson="plot-options", fig.cap="Proportional change in species' mean cover caused by a 1% increase in observed precipitation, temperature, or both as predicted by the individual-based IPM and the aggregate-based QBM. Note that the change for POSE due to a precipitation increase predicted by the QBM is almost zero and so does not show up in the figure. Labels on x-axis refer to: 'ppt' = 1% increase in mean precipitation; 'temp' = 1% increase in mean temperature; 'ppt + temp' = 1% increase in mean precipitation and mean temperature.", fig.width=8.5, fig.height=3, cache=FALSE, results='hide'}
ipm_diffs <- readRDS("../montana/ipm/simulations/results/ipm_climatesims_logdiffs.RDS")
ipm_diffs$model <- rep("IPM", nrow(ipm_diffs))
ipm_diffs$type <- rep("paramunc", nrow(ipm_diffs))
ipm_diffs2 <- readRDS("../montana/ipm/simulations/results/ipm_climatesims_logdiffs_mean.RDS")
ipm_diffs2$model <- rep("IPM", nrow(ipm_diffs))
ipm_diffs2$type <- rep("parammean", nrow(ipm_diffs))
ipm_diffs$med_cover <- ipm_diffs2$med_cover
ipm_diffs <- rbind(ipm_diffs, ipm_diffs2)

>>>>

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
bibliography: ~/Dropbox/Bibliography/MicroMeso.bib
csl: components/ecology.csl

@@

Our analysis offers good news and bad news. The good news is that for two out of our four focal species, the QBM faired no worse than the demographically detailed IPM in terms of mean absolute (Table 1). Thus, it seems that aggregate level data can provide forecasts of similar skill as individual level data.

@@

##  QBM results
# Read in data
qbm_onestep <- readRDS("../analysis/quadBM/simulations/validation_results/qbm_one-step_forecasts_combined.rds")
# Average predictions over quad-year reps
avg_qbm <- ddply(qbm_onestep, .(species, quad, year), summarise,
                 avg_prediction = median(cover.t1),
                 observation = mean(cover.t0),
                 quant_dist = quantile(cover.t1*100, probs = 0.95) - quantile(cover.t1, probs = 0.05))
# Calculate error
avg_qbm$error <- avg_qbm$observation*100 - avg_qbm$avg_prediction*100
# Aggregate over species
qbmstats <- ddply(avg_qbm, .(species), summarise,
                  mae = mean(abs(error)),
                  mean_cover = mean(observation*100),
                  quant_dist = mean(quant_dist))
qbmstats$model <- "QBM"

# mae_list <- list()
# mae_pvals <- list()
# species <- unique(qbmstats$species)
# for(spp in species){
#   tmp1 <- subset(tmpipm, species==spp)
#   tmp2 <- subset(tmpqbm, species==spp)
#   err1 <- as.data.frame(abs(tmp1$mean_obs - tmp1$mean_pred))
#   err2 <- as.data.frame(abs(tmp2$mean_obs - tmp2$mean_pred))
#   colnames(err1) <- "resid"
#   err1$model <- "ipm"
#   colnames(err2) <- "resid"
#   err2$model <- "qbm"
#   errd <-  rbind(err1, err2)
#   mae_ttest <- with(errd, t.test(resid~model, alternative = "less"))
#   mae_list[[spp]] <- mae_ttest
#   mae_pvals[[spp]] <- mae_ttest[["p.value"]]
# }
# 
# mae_ps <- melt(mae_pvals)

@@

\pagebreak{}
```{r table_2,results='asis',message=FALSE, comment=NA, echo=FALSE}
# qbm_diffs <- readRDS("../montana/quadBM/simulations/results/qbm_climatesims_logdiffs.RDS")
# qbm_diffs$type <- rep("QBM", nrow(qbm_diffs))
# ipm_diffs <- readRDS("../montana/ipm/simulations/results/ipm_climatesims_logdiffs.RDS")
# ipm_diffs$type <- rep("IPM", nrow(ipm_diffs))
# 
# diff_df <- rbind(qbm_diffs, ipm_diffs)
# med_cover_perc <- as.data.frame(diff_df$med_cover*100)
# med_cover_perc$model <- diff_df$type 
# med_cover_perc$sim <- diff_df$variable
# med_cover_perc$species <- diff_df$species
# colnames(med_cover_perc)[1] <- "cover"
# medcast <- dcast(med_cover_perc, species+sim~model, value.var = "cover")
# colnames(medcast) <- c("Species", "Simulation", "IPM Change (%)", "QBM Change (%)")
# simnames <- c("ppt", "temp", "ppt+temp")
# medcast$Simulation <- rep(simnames, 4)
# print.xtable(xtable(medcast, caption = "Percent change from equilibrium (temporal median of simulated plant cover) at observed climate relative to a 1\\% climate change as forecast by each model type."),type="latex", comment=FALSE,
#              include.rownames=FALSE, caption.placement="top")
```

@@

Figures
-------
```{r figure_1,dependson="plot-options", fig.cap="Work flow of the data aggregation, model fitting, and population simulating.", fig.height=6}
library(png)
library(grid)
img <- readPNG("components/figure/micromeso_flow.png")
grid.raster(img)
```

@@

```{r figure_2, dependson="plot-options", fig.cap="Sensitivity of equilibrium cover to a 1% increase in precipitation (+ppt), temperature (+temp), or both (+ppt&temp) applied to individual and combined vital rates. For example, the points associated with G show the median cover from IPM simulations where a climate perturbation is applied only to the growth regression climate covariates.", fig.width=4.5, fig.height=6, cache=FALSE, results='hide'}
setwd("../analysis/ipm/simulations/results/climate_sensitivity/")
files <- list.files()
cover_files <- files[grep("cover", files)]
spp_id <- substr(cover_files, 1, 4)
num_files <- length(cover_files)
all_sims <- data.frame(time=NA, cover=NA, species=NA, climsim=NA, vital=NA)
for(i in 1:num_files){
  tmp <- read.csv(cover_files[i])
  tmp$species <- rep(spp_id[i], nrow(tmp))
  all_sims <- rbind(all_sims, tmp)
}
all_sims <- all_sims[2:nrow(all_sims),]

@@

equilibrium_cover <- subset(equilibrium_cover, vital!="A")
bothid <- which(equilibrium_cover[,"climsim"]=="ppttemp")
equilibrium_cover[bothid,"climsim"] <- "zppttemp"
# myCols2 <- c("#277BA8", "#7ABBBD", "#AED77A")
ggplot(equilibrium_cover, aes(x=climsim, y=eq_cover*100, 
                              shape=vital, group=vital, linetype=vital))+
  geom_hline(data=mean_cover, aes(yintercept=value*100), linetype=2, color="grey45")+
  geom_line()+
  geom_point(size=4)+
  facet_grid(species~., scales = "free")+
  scale_linetype_discrete(name="Vital Rate")+
  scale_shape_discrete(name="Vital Rate")+
  scale_x_discrete(labels=c("baseline", "+ppt", "+temp", "+ppt&temp"))+
  ylab("Equilibrium Cover (%)")+
  xlab("Simulation")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

covertest <- subset(equilibrium_cover, climsim=="all")
colnames(covertest)[4] <- "obs_cover"
out <- merge(subset(equilibrium_cover, climsim!="all"), covertest[,c("species","vital","obs_cover")])
out$propdiff <- with(out, log(eq_cover*100)-log(obs_cover*100))
setwd("~/Repos/MicroMesoForecast/manuscript")
```

```{r figure_3, dependson="plot-options", fig.cap="Proportional change in species' mean cover caused by a 1% increase in observed precipitation, temperature, or both as predicted by the individual-based IPM and the aggregate-based QBM. Note that the change for POSE due to a precipitation increase predicted by the QBM is almost zero and so does not show up in the figure. Labels on x-axis refer to: 'ppt' = 1% increase in mean precipitation; 'temp' = 1% increase in mean temperature; 'ppt + temp' = 1% increase in mean precipitation and mean temperature.", fig.width=8.5, fig.height=3, cache=FALSE, results='hide'}
# ipm_diffs <- readRDS("../montana/ipm/simulations/results/ipm_climatesims_logdiffs.RDS")
# ipm_diffs$model <- rep("IPM", nrow(ipm_diffs))
# ipm_diffs$type <- rep("paramunc", nrow(ipm_diffs))
# ipm_diffs2 <- readRDS("../montana/ipm/simulations/results/ipm_climatesims_logdiffs_mean.RDS")
# ipm_diffs2$model <- rep("IPM", nrow(ipm_diffs))
# ipm_diffs2$type <- rep("parammean", nrow(ipm_diffs))
# ipm_diffs$med_cover <- ipm_diffs2$med_cover
# ipm_diffs <- rbind(ipm_diffs, ipm_diffs2)

### 2015-06-15 manuscript/manuscript.tex
adds QBM loyo simulation results and updates ms Table 1 with those results

\author{}
\date{}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\doublespacing

@@

\begin{table}[ht]
\centering
\caption{Mean absolute error (MAE) and accuracy (Pearson's $\rho$) for one-step-ahead forecast from both model types. Forecasts were made without random year effects; only climate covariates could explain year-to-year variation. 90 Distance refers to the average distance between the upper and lower 90 percentiles of the 100 predicted values for each quadrat-year combination.} 
\begin{tabular}{llrrr}
  \hline
Species & Model & MAE & 90\% Distance & Mean Obs. Cover \\ 
  \hline
BOGR & IPM & 4.77 & 29.33 & 9.16 \\ 
  BOGR & QBM & 5.30 & 44.63 & 9.22 \\ 
  HECO & IPM & 0.60 & 2.82 & 1.22 \\ 
  HECO & QBM & 2.03 & 12.05 & 1.25 \\ 
  PASM & IPM & 0.18 & 0.49 & 0.40 \\ 
  PASM & QBM & 0.17 & 1.34 & 0.40 \\ 
  POSE & IPM & 0.78 & 2.16 & 1.23 \\ 
  POSE & QBM & 1.30 & 7.04 & 1.25 \\ 
   \hline
\end{tabular}
\end{table}

@@

\begin{table}[ht]
\centering
\caption{Percent change from equilibrium (temporal median of simulated plant cover) at observed climate relative to a 1\% climate change as forecast by each model type.} 
\begin{tabular}{llrr}
  \hline
Species & Simulation & IPM Change (\%) & QBM Change (\%) \\ 
  \hline
BOGR & ppt & 0.92 & 19.43 \\ 
  BOGR & temp & 12.50 & -93.56 \\ 
  BOGR & ppt+temp & 25.06 & -27.88 \\ 
  HECO & ppt & -6.65 & -9.40 \\ 
  HECO & temp & -6.46 & -10.63 \\ 
  HECO & ppt+temp & -25.72 & -14.27 \\ 
  PASM & ppt & -8.53 & 0.25 \\ 
  PASM & temp & 8.03 & 8.06 \\ 
  PASM & ppt+temp & 8.74 & 1.30 \\ 
  POSE & ppt & 1.71 & -8.99 \\ 
  POSE & temp & 20.06 & -16.80 \\ 
  POSE & ppt+temp & 11.06 & 4.04 \\ 
   \hline
\end{tabular}
\end{table}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-1.pdf}
\caption{Work flow of the data aggregation, model fitting, and
population simulating.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_1.pdf}
\caption{Sensitivity of equilibrium cover to removal of climate effects
from each vital rate regression.}
\end{figure}

>>>>

\author{}
\date{}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\doublespacing
\setlength\parindent{24pt}

@@

\begin{table}[ht]
\centering
\caption{Mean absolute error (MAE) and accuracy (Pearson's $\rho$) for one-step-ahead forecast from both model types. Forecasts were made without random year effects; only climate covariates could explain year-to-year variation. 90 Distance refers to the average distance between the upper and lower 90 percentiles of the 100 predicted values for each quadrat-year combination.} 
\begin{tabular}{llrrr}
  \hline
Species & Model & MAE & 90\% Distance & Mean Obs. Cover \\ 
  \hline
BOGR & IPM & 12.18 & 38.52 & 9.43 \\ 
  BOGR & QBM & 19.66 & 56.50 & 9.26 \\ 
  HECO & IPM & 1.22 & 6.47 & 1.15 \\ 
  HECO & QBM & 12.35 & 41.11 & 1.18 \\ 
  PASM & IPM & 0.19 & 1.65 & 0.42 \\ 
  PASM & QBM & 0.55 & 7.78 & 0.42 \\ 
  POSE & IPM & 1.37 & 7.64 & 1.25 \\ 
  POSE & QBM & 1.79 & 40.59 & 1.27 \\ 
   \hline
\end{tabular}
\end{table}

@@

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_1.pdf}
\caption{Work flow of the data aggregation, model fitting, and
population simulating.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_2.pdf}
\caption{Sensitivity of equilibrium cover to a 1\% increase in
precipitation (+ppt), temperature (+temp), or both (+ppt\&temp) applied
to individual and combined vital rates. For example, the points
associated with G show the median cover from IPM simulations where a
climate perturbation is applied only to the growth regression climate
covariates.}
\end{figure}
